---
title: "Raising the Curtain: Getting Started with overtureR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping with OvertureR}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10, 
  fig.height = 6,
  out.width = "100%",
  comment = "#>"
)

```

```{r eval=FALSE}
# install if needed:
install.packages("overtureR")
```

```{r setup}
library(overtureR)
library(ggplot2)
library(dplyr)
library(sf)
```

This vignette demonstrates how to use overtureR to access and visualize Overture Maps data, focusing on a practical example in Washington, DC: finding the theater.

Overture Maps is an open-source mapping initiative aimed at developers who build map services or use geospatial data. It provides a collaborative, globally-referenced, and quality-assured dataset with a structured schema. This makes it an excellent resource for creating reliable and interoperable map products.
Using overtureR, we can easily tap into this rich dataset. In this guide, we'll walk through the process of:

1. Fetching the boundary of Washington, DC
2. Locating Ronald Reagan National Airport
3. Finding the Kennedy Center theater
4. Getting to the Kennedy Center with public transit

`open_curtain()` function is our primary tool for accessing Overture Maps data. We'll start by using `open_curtain()` to retrieve the DC boundary and pinpoint the airport:

```{r reagan}
# Washington, DC boundary
dc <- open_curtain("division_area") |> 
  filter(subtype == "region", region == "US-DC") |> 
  collect()

# adding a bounding box makes the query faster:
dc_catchment <- st_geometry(dc) |>
  # 10 miles from DC
  st_buffer(10 * 1609.34) |> 
  st_bbox()

reagan_airport <- open_curtain("place", spatial_filter = dc_catchment) |> 
    filter(
      names$primary == "Ronald Reagan Washington National Airport",
      categories$primary == "airport"
    ) |> 
    collect()

print(reagan_airport)
```

By default, `open_curtain` would search through every "place" (aka point of interest) in the world - an enormous dataset. Obviously, that's too much to load into most computers' memory, so `open_curtain` does this lazily. Only after calling `collect_sf` does it load data onto your computer. So we filter the data first, spatially and by name, like so:

1. fetch the boundary of Washington, DC from the "division_area" dataset;
2. filter for the specific region we wanted;
3. create a spatial buffer around DC to define our area of interest for subsequent queries; and
4. locate Ronald Reagan National Airport using the "place" dataset, filtering by name and category.

Afterwards, `collect_sf` brings the only the data need into memory. For more on lazy programming, see the [dbplyr documentation](https://dbplyr.tidyverse.org/).

Now that we've set the stage with our starting point, let's spotlight our destination. In the next code block, we'll locate the Kennedy Center:

```{r reagan_plot}
reagan_plot <- ggplot() + 
  geom_sf(data = dc, fill = "purple", alpha = 0.05) + 
  geom_sf(data = reagan_airport, color = "red", size = 4) +
  geom_sf_label(
    data = reagan_airport, nudge_y = 0.01, aes(label = names$primary)
  ) +
  theme_minimal() +
  theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank()
    )

reagan_plot
```

In this code, we've queried the "building" dataset within our defined DC area. We used a text filter to find buildings with "The Kennedy Center" in their name. This demonstrates overtureR's ability to perform text-based searches within the Overture Maps dataset.

To get to the theater, we'll need to know our transit options. The following code showcases overtureR's capacity to handle more complex spatial and attribute queries:

```{r kennedy}
kennedy_center <- open_curtain("building", st_bbox(dc)) |>
  filter(grepl("Kennedy Center", names$primary)) |> 
  collect()

kennedy_plot <- reagan_plot + 
  geom_sf(data = kennedy_center, fill = "green") +
  geom_sf_label(data = kennedy_center, nudge_y = 0.01, aes(label = names$primary)) 
kennedy_plot
```

In the code above, we've created a bounding box that encompasses both the airport and the Kennedy Center, plus a one-mile buffer. We then used this to filter the "segment" dataset for rail transit, specifically the Blue Line of the DC Metro.

For the grand finale, we'll create a map that displays all the elements we've gathered:

```{r metro}
# filter town to areas that are within 1 mile of our two points
kennedy_reagan_bbox <- bind_rows(kennedy_center, reagan_airport) |> 
  st_bbox() |> 
  st_as_sfc() |> 
  st_buffer(1 * 1609.34) |> 
  st_bbox()

dc_transit <- open_curtain("segment", kennedy_reagan_bbox) |> 
  filter(
    subtype == "rail", 
    # filter to the Blue Line of the DC Metro
    grepl("Metro", names$primary),
    grepl("Blue", names$primary)
  ) |> 
  select(id, names, geometry) |> 
  collect()

print(dc_transit)
```

This final step uses ggplot2 to create a map that displays the airport, the Kennedy Center, and the Metro Blue Line connecting them. This visualizes the route from our arrival point to our theatrical destination.

```{r kennedy_plot}

kennedy_plot +
  geom_sf(data = dc_transit, color = "blue") +
  coord_sf(
    xlim = c(kennedy_reagan_bbox[["xmin"]], kennedy_reagan_bbox[["xmax"]]),
    ylim = c(kennedy_reagan_bbox[["ymin"]], kennedy_reagan_bbox[["ymax"]]),
  )

```


Perfect, it looks like we can take the blue line straight there. Break a leg!
